---
title: "A Brief Overview"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(plotly)
library(gapminder)
library(DT)
library(htmlwidgets)
library(base)
library(hablar)
library(tidyr)
library(viridis)
library(hrbrthemes)
```

The plot below is an approximate density histogram for the number of citizens killed by each of the 100 police departments, broken down by race from 2013-2019 (inclusive). A quick look shows us that, in general, BAME citizens are killed in larger numbers across the 100 departments compared to white citizens.

```{r Deaths Overview,echo = FALSE,fig.align='center'}
FinalData <- read.csv("~/OneDrive/Documents/Kaggle/697683_1239276_bundle_archive/FinalData.csv",header = TRUE)
df <- rbind(data.frame(x = FinalData$White.deaths, Race ='White'),data.frame(x = FinalData$BAME.deaths, Race ='BAME'))
ggplot(df, aes(x, group=Race, col=Race)) + geom_density(position=position_dodge(width = 0)) + xlab('Deaths') + ylab("Density") 
```

Perhaps this is a little misleading since this plot hasn't been scaled by population size. A [US Census](https://www.census.gov/quickfacts/fact/table/US/RHI125218) estimated that 60.4% of the US population are 'White alone, not Hispanic nor Latino'. In other words, even if we scaled this by number of citizens of each race, say deaths per hundred thousand people, the characteristics of the above plot are only amplified. 

<br>

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>
<font size = "5">
<p class="comment">
"Police contacted Henry Lane shortly after midnight for unspecified reasons. Red Bluff Cpl. Stephen Harper shot and killed him when Lane threatened officers with a (walking) stick."
</p>
</font>


