---
title: "Deaths Per Capita"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(plotly)
library(gapminder)
library(DT)
library(htmlwidgets)
library(base)
library(hablar)
library(tidyr)
library(viridis)
library(hrbrthemes)
```

The below plot shows a transformation of the number of White deaths per capita (White population) vs the number of BAME deaths per capita (BAME population) within each police institution. The black line represents an equal death rate in which police kill citizens of each race at equal rates.

```{r Deaths per Captia,echo = FALSE,fig.align='center'}
FinalData <- mutate(FinalData,BAME = FinalData$Black + FinalData$Hispanic + FinalData$Asian + FinalData$Amer..Indian + FinalData$Hawaiian + FinalData$Two.or.more.races)
FinalData <- mutate(FinalData,White.Deaths.Per.Capita = FinalData$White.deaths/FinalData$White)
FinalData <- mutate(FinalData,BAME.Deaths.Per.Capita = FinalData$BAME.deaths/FinalData$BAME)

df <- data.frame(FinalData$White.Deaths.Per.Capita,FinalData$BAME.Deaths.Per.Capita)
df <- data.frame(df,FinalData$Total)
df <- data.frame(df,FinalData$PD)

p <- ggplot(df,aes(log10(FinalData.White.Deaths.Per.Capita+1),log10(FinalData.BAME.Deaths.Per.Capita+1),size = FinalData.Total,color = "red",text = paste("",FinalData.PD,"<br>Population:",FinalData.Total))) + geom_point(alpha = 0.5)+xlim(0,8e-05)+scale_size(range = c(.1, 12), name="Population (M)")+theme(legend.position="none")+xlab("log(White Deaths Per Capita + 1)")+ylab("log(BAME Deaths Per Capita + 1)")+coord_equal() + geom_abline(slope=1, intercept=0) + labs(title = "Citizen Deaths by Police Department")
ggplotly(p,tooltip = "text")
```

Clearly the vast majority of the data points lie above the equal rates line.

<br><br><br><br><br><br><br><br><br><br><br><br>


<font size = "2">

##### Justifying the log transformation
If we made the assumption that there was a linear relationship between the variables, we would be interested in the value of the gradient of the line of best fit. This straight line will have the form: $log_{10}(BDPC+1)=alog_{10}(WDPC+1)+b$ or equivalently $BDPC=(WDPC+1)^{a}+10^{b}-1$. For small WDPC, we can use the expansion $(1+x)^{a}\approx1+ax$ for small $x$ to give $BDPC\approx 1+a(WDPC)+10^{b}-1=a(WDPC)+10^{b}$. So, for small values of deaths per capita, this means that the log transformation has made the plot more clear, but hasn't changed the meaning of parameter $a$. In either case, it effectively scales deaths per capita of White citizens to BAME citizens. One can visibly see that $a>1$, which is sufficient to claim that BAME citizens are killed more frequently than White citizens, when scaled by population.

</font>