---
title: "Deaths Per Capita"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
library(plotly)
library(gapminder)
library(DT)
library(htmlwidgets)
library(base)
library(hablar)
library(tidyr)
library(viridis)
library(hrbrthemes)
```

The below plot shows a transformation of the number of White deaths per capita (White population) vs the number of BAME deaths per capita (BAME population) within each police institution. The black line represents an equal death rate in which police kill citizens of each race at equal rates. The log transformation has been used to make the plot look clearer but without changing its interpretation.

```{r Deaths per Captia,echo = FALSE,fig.align='center',fig.width=9,fig.height=8}
FinalData <- mutate(FinalData,BAME = FinalData$Black + FinalData$Hispanic + FinalData$Asian + FinalData$Amer..Indian + FinalData$Hawaiian + FinalData$Two.or.more.races)
FinalData <- mutate(FinalData,White.Deaths.Per.Capita = FinalData$White.deaths/FinalData$White)
FinalData <- mutate(FinalData,BAME.Deaths.Per.Capita = FinalData$BAME.deaths/FinalData$BAME)

df <- data.frame(FinalData$White.Deaths.Per.Capita,FinalData$BAME.Deaths.Per.Capita)
df <- data.frame(df,FinalData$Total)
df <- data.frame(df,FinalData$PD)

p <- ggplot(df,aes(log10(FinalData.White.Deaths.Per.Capita+1),log10(FinalData.BAME.Deaths.Per.Capita+1),size = FinalData.Total,color = "red",text = paste("",FinalData.PD,"<br>Population:",FinalData.Total))) + geom_point(alpha = 0.5)+xlim(0,8e-05)+scale_size(range = c(.1, 12), name="Population (M)")+theme(legend.position="none")+xlab("log(White Deaths Per Capita + 1)")+ylab("log(BAME Deaths Per Capita + 1)")+coord_equal() + geom_abline(slope=1, intercept=0) + labs(title = "Citizen Deaths by Police Department")
ggplotly(p,tooltip = "text")
```

Clearly the vast majority of the data points lie above the equal rates line- BAME people are killed at a higher rate than White people across the vast majority of departments, when scaled by population size.

<br>

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>
<font size = "5">
<p class="comment">
"Euharlee police said 17-year-old Christopher Roupe pointed a gun at one of their officers on Feb. 14, when officers knocked on the door of the family's home to serve a probation warrant for Roupe's father. Family members, however, said the teen was holding a Nintendo Wii controller."
</font>
</p>
