---
title: "Motivation"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rmarkdown)
library(ggplot2)
library(dplyr)
library(plotly)
library(gapminder)
library(DT)
library(htmlwidgets)
library(base)
library(hablar)
library(tidyr)
library(viridis)
library(hrbrthemes)
```

Catalysed by the killing of George Floyd, police departments across the US have come under public scrutiny. This was just one of the many recorded examples of police brutality against BAME citizens which, for many, was the final straw in remaining compliant. In light of this, data collected mainly from [Washington Post](https://github.com/washingtonpost/data-police-shootings), [FiveThirtyEight](https://projects.fivethirtyeight.com/republicans-democrats-cities/) and US Census Data may be used to visualise the extent of police brutality and also the disparity in treatment of citizens of different races.

<br>


## The Data

A brief snippet of the data can be seen below. There are only 200 rows shown, but the data set actually covers the deaths of 7907 people. Some of the more important columns have been included in the interest of brevity, however there is much more information available.

```{r Police Killings Data,echo = FALSE}
police_killings <- read.csv("~/OneDrive/Documents/Kaggle/697683_1239276_bundle_archive/police_killings.csv",header = TRUE)
police_killings2 <- police_killings[c(1:200),c(1,2,3,4,8,12,13,14,19,20)]
datatable(police_killings2, options = list(pageLength = 5,pageWidth=5),rownames = FALSE,colnames = c("Victim","Age","Sex","Race","City","Responsible Police Agency","Cause of Death","Brief Description","Armed?","Alleged Weapon"))
```

<br>

<style>
p.comment {
background-color: #DBDBDB;
padding: 10px;
border: 1px solid black;
margin-left: 25px;
border-radius: 5px;
font-style: italic;
}

</style>
<font size = "5">
<p class="comment">
"Police said they shot and killed the man after he hit the windshield of a patrol car and ignored commands from police. He was carrying a toy broomstick."
</p>
</font>



